---
title: "HW2_cs3652"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

```{r}
library(tidyverse)
```

##Problem 1

```{r, message = FALSE}
subway_data = read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", col_types = "cccddcccccccccccccccccccccccddcc") %>%
  janitor::clean_names() %>% 
  ##cleaning names so that they are lowercase snake
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  ##selecting relevant variables
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE))
  ##recoding to logical variable 
```
  
This dataset has many variables that describe the MTA service at New York City subway stations. It includes the routes that each station serves, as well as entry/exit points, and ada compliance, among many other things. Thus far, the data cleaning steps that I have taken so far are: cleaning the names so that they are easy to read and are lower case snake, selecting the columns (variables) that are relevant to our analysis, and mutating the entry variable so that it is a logical variable. The dimension of the resulting dataset is `r dim(subway_data)` (that is 1868 rows x 19 columns). These data are not entirely tidy since the route variables can be futher simplified into more usable data based on the analyses that we want to conduct. 

```{r, message = FALSE}
subway_data %>% 
  distinct(line, station_name) 
  ##finding the distinct number of stations 
select(subway_data, line, station_name, ada) %>% 
  filter(ada == "TRUE") %>% 
  distinct(line, station_name, ada)
  ##finding the number of ADA compliant stations 
nrow(filter(subway_data, entry == TRUE, vending == "NO")) / nrow(filter(subway_data, vending == "NO"))
  ##finding the proportion of station entrances without vending allowing entrance. 
```

According to this dataset there are `r count(distinct(subway_data, line, station_name))` distinct stations. There are `r nrow(filter(distinct(subway_data, line, station_name, ada), ada == TRUE))` stations that are ADA compliant. Furthermore, the proportion of stations extrances/exits, without vending, that allow entrance is `r nrow(filter(subway_data, entry == TRUE, vending == "NO"))/nrow(filter(subway_data, vending == "NO"))` or 37.7%.  

```{r, message = FALSE}
subway_data_tidy = (gather(subway_data, key = route_number, value = route, route1:route11)) 
#creating a new dataframe where route number information is simplified
```

```{r, message = FALSE}
nrow(filter(distinct(subway_data_tidy, line, station_name, route), route == "A"))

nrow(filter(distinct(subway_data_tidy, line, station_name, route, ada), route == "A", ada == TRUE))
```

There are `r nrow(filter(distinct(subway_data_tidy, line, station_name, route), route == "A"))` stations on the A line. Only `r nrow(filter(distinct(subway_data_tidy, line, station_name, route, ada), route == "A", ada == TRUE))` of the 60 stations are ADA compliant.

Problem 2

This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website. Please use the HealthyHarborWaterWheelTotals2017-9-26.xlsx version.

Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
    use reasonable variable names
    omit rows that do not include dumpster-specific data
    rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016?
Problem 3

This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package.

For this question:

format the data to use appropriate variable names;
    focus on the “Overall Health” topic
    exclude variables for class, topic, question, sample size, and everything from lower confidence limit to GeoLocation
    structure data so that values for Response (“Excellent” to “Poor”) are column names / variables which indicate the proportion of subjects with each response (which are values of Data_value in the original dataset)
    create a new variable showing the proportion of responses that were “Excellent” or “Very Good”

Using this dataset, do or answer the following:

How many unique locations are included in the dataset? Is every state represented? What state is observed the most?
    In 2002, what is the median of the “Excellent” response value?
    Make a histogram of “Excellent” response values in the year 2002.
    Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.


